---
title: "Logistic Regression"
author: "DS 501 - Introduction to Data Science"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r, message=FALSE, echo=FALSE}
setwd(getwd())
```

### Loading packages that we will be using
```{r, message=FALSE}
library(ggplot2)
library(ROCR)
library(popbio)
library(readr)
library(superml)
library(dplyr)
```

### Load and explore the data
```{r, message=FALSE}
stroke <- read_csv("stroke.csv")
head(stroke)
strokeData = stroke[,2:12]
head(strokeData)
```

### How many patients had stroke?
```{r}
sum(strokeData$stroke)
```

### Data cleanup
```{r}
strokeData[strokeData == "N/A"] <- NA
cleanStrokeData <- na.omit(strokeData)

cleanStrokeData$gender <- as.factor(cleanStrokeData$gender)
cleanStrokeData$hypertension <- as.factor(cleanStrokeData$hypertension)
cleanStrokeData$heart_disease <- as.factor(cleanStrokeData$heart_disease)
cleanStrokeData$ever_married <- as.factor(cleanStrokeData$ever_married)
cleanStrokeData$work_type <- as.factor(cleanStrokeData$work_type)
cleanStrokeData$Residence_type <- as.factor(cleanStrokeData$Residence_type)
cleanStrokeData$smoking_status <- as.factor(cleanStrokeData$smoking_status)
cleanStrokeData$bmi <- as.numeric(cleanStrokeData$bmi)
```
### Plot the data
```{r}
 qplot(bmi, stroke, data=cleanStrokeData) + geom_point(colour = "#3366FF", size = 3)
```

### Plot the conditional denisty plot
Computes and plots conditional densities describing how the conditional distribution of a categorical variable `stroke` changes over a numerical variable `age`.
```{r}
cdplot(factor(stroke) ~ cleanStrokeData$avg_glucose_level, data=cleanStrokeData, main="Estimated categ prob", ylab='Stroke')
qplot(avg_glucose_level, ..count.., data=cleanStrokeData, geom="density", fill=factor(stroke), position="fill") + 
  ylab('Probability')+theme(legend.position='bottom')

```

```{r}
cdplot(factor(stroke) ~ cleanStrokeData$age, data=cleanStrokeData, main="Estimated categ prob", ylab='Stroke')
qplot(age, ..count.., data=cleanStrokeData, geom="density", fill=factor(stroke), position="fill") + 
  ylab('Probability')+theme(legend.position='bottom')

```

```{r}
cdplot(factor(stroke) ~ cleanStrokeData$bmi, data=cleanStrokeData, main="Estimated categ prob", ylab='Stroke')
qplot(bmi, ..count.., data=cleanStrokeData, geom="density", fill=factor(stroke), position="fill") + 
  ylab('Probability')+theme(legend.position='bottom')

```


```{r}

ggplot(cleanStrokeData
       , aes(bmi, stroke))+
  geom_point() +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family="binomial"))



```


### Perform the logistic regression
Function `glm` is used to fit generalized linear models, specified by giving a symbolic description of the linear predictor and a description of the error distribution.

```{r}
strokeLogreg = glm(stroke ~ hypertension+heart_disease+avg_glucose_level+bmi, data=cleanStrokeData, family=binomial(link="logit"))
summary(strokeLogreg)
```

### Generate confidence intervals for regression coefficients
```{r}
confint(strokeLogreg)
```

### Split the data as Training and Test sets
```{r}
set.seed(2015)
## Convert stroke to factor
cleanStrokeData <- cleanStrokeData %>% mutate(stroke = as.factor(cleanStrokeData$stroke))
## Generate random ordering
rand <- sample(nrow(cleanStrokeData))
cleanStrokeData <- cleanStrokeData[rand,]
splitStroke = caret::createDataPartition(cleanStrokeData$stroke, p = 0.8, list=F, times=1)
head(splitStroke)
trainStroke = cleanStrokeData[splitStroke,]
head(trainStroke)
# testStroke = cleanStrokeData[-!row.names(strokeData) %in% row.names(trainStroke),]
testStroke = cleanStrokeData[-splitStroke,]
head(testStroke)
```

### Apply Logistic Regression on Training set
```{r}
trainStrokeLR = glm(stroke ~ age+hypertension+heart_disease+avg_glucose_level+bmi, data=trainStroke, family=binomial(link="logit"))
summary(trainStrokeLR)
```

### Predict on Test data set
```{r}
testStroke$Predicted = round(predict(trainStrokeLR,
                                     testStroke[,c( 2, 3, 4, 8, 9)],
                                     type="response"), 2)
head(testStroke)
```

### Visualization of Logistic Regression results
* `popbio` package
```{r}
plot(trainStroke$bmi, trainStroke$stroke, xlab="bmi", ylab="P(stroke)")
trainLR = glm(stroke ~ bmi, data=trainStroke, family=binomial(link="logit"))
curve(predict(trainLR,data.frame(bmi=x),type="resp"),add=TRUE, n=10000, type = "l")
points(trainStroke$bmi,fitted(trainLR),pch=20)
popbio::logi.hist.plot(as.numeric(trainStroke$bmi), trainStroke$stroke, logi.mod = 1, boxp=FALSE,type="hist" )
```


```{r}
trainStroke %>%
#  mutate(prob = ifelse(diabetes == "pos", 1, 0)) %>%
  ggplot(aes(bmi, stroke)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  labs(
    title = "Logistic Regression Model", 
    x = "Body Mass Index",
    y = "Probability of stroke"
    )

```






* `visreg` package
```{r}
logReg = glm(stroke ~ bmi + gender, data=trainStroke, family=binomial(link="logit"))
visreg::visreg(logReg, "bmi", scale="response", partial=FALSE, xlab="bmi", ylab="P(stroke)", rug=2, by = "gender")
```
```{r}
logReg = glm(stroke ~ age, data=trainStroke, family=binomial(link="logit"))
visreg::visreg(logReg, "age", scale="response", partial=FALSE, xlab="age", ylab="P(stroke)", rug=2)
```


```{r}
logReg = glm(stroke ~ avg_glucose_level + gender, data=trainStroke, family=binomial(link="logit"))
visreg::visreg(logReg, "avg_glucose_level", scale="response", partial=FALSE, xlab="avg_glucose_level", ylab="P(stroke)", rug=2, by ="gender")
```


```{r}
logReg = glm(stroke ~ avg_glucose_level, data=trainStroke, family=binomial(link="logit"))
visreg::visreg(logReg, "avg_glucose_level", scale="response", partial=FALSE, xlab="avg_glucose_level", ylab="P(stroke)", rug=2)
```





### Model evaluation - Receiver Operating Characteristic (ROC) Curve
Every classifier evaluation using ROCR starts with creating a prediction object. This function is used to transform the input data into a standardized format.

```{r}
pred = predict(trainStrokeLR, testStroke[,c(2, 3, 4, 8, 9)], type="response")
pObject = ROCR::prediction(pred, testStroke$stroke)
```

All kinds of predictor evaluations are performed using `performance` function

```{r}
rocObj = ROCR::performance(pObject, measure="tpr", x.measure="fpr")
aucObj = ROCR::performance(pObject, measure="auc")  
plot(rocObj, main = paste("Area under the curve:", round(aucObj@y.values[[1]] ,4))) 
```

### Random Model 
```{r}
trainStrokeRandom = trainStroke
set.seed(1235)
trainStrokeRandom$stroke = sample(c(0,1), replace=T, size=nrow(trainStroke))
logRegRandom = glm(stroke ~ hypertension+heart_disease+avg_glucose_level+bmi, data=trainStrokeRandom, family=binomial(link="logit"))

rand_pred = predict(logRegRandom, testStroke[,c(3, 4, 8, 9)], type="response")
randObject = ROCR::prediction(rand_pred, testStroke$stroke )
rocRandObj = ROCR::performance(randObject, measure="tpr", x.measure="fpr")
aucRandObj = ROCR::performance(randObject, measure="auc")  
plot(rocRandObj, main = paste("Area under the curve:", round(aucRandObj@y.values[[1]] ,4))) 
```

### Assess model fit
```{r}
Phat = predict(trainStrokeLR,testStroke,type="response")
head(Phat)
prop.table(xtabs(~ stroke, data=testStroke))
thresh = 0.5
facHat = cut(Phat, breaks=c(-Inf, thresh, Inf), labels=c(0, 1))
cTab   = xtabs(~ stroke + facHat, data=testStroke)
addmargins(cTab)
```

* Correct classification rate
```{r}
CCR = sum(diag(cTab)) / sum(cTab)
CCR
```

### Updating models
If you want to modify a model you may consider using the special function `update`
```{r, eval=FALSE}
lrUpdate = update(trainChurnLR, ~ . -Married-Churned_contacts-Cust_years)
summary(lrUpdate)
pred = predict(lrUpdate, testChurn[,2:5], type="response")
pObject = ROCR::prediction(pred, testChurn$Churned )
rocObj = ROCR::performance(pObject, measure="tpr", x.measure="fpr")
aucObj = ROCR::performance(pObject, measure="auc")  
plot(rocObj, main = paste("Area under the curve:", round(aucObj@y.values[[1]] ,4))) 
```

